### 2017-04-11 paper/oopsla17-submission/evaluation-utility.tex
prune some of the more obnoxious orphans

\subsubsection{Type Error Slice}
\label{sec:type-error-slice}
First we must justify our decision to automatically exclude expressions
outside the minimal type error slice from consideration.
%
% The \InSlice feature should be highly predictive --- a fix must change
% at least one expression in the type error slice.
% %
% Thus, our first experiment seeks to quantify the impact of \InSlice by
% comparing the accuracy of our classifiers on three sets of features:
%
Thus, we compare the accuracy of our classifiers on three sets of features:
%
\begin{enumerate}
\item A baseline with only local syntactic features and no
  preemptive filtering by \InSlice.
\item The features of (1) extended with \InSlice.
\item The same features as (1), but we preemptively discard samples
  where \InSlice is disabled.
\end{enumerate}
%
The key difference between (2) and (3) is that a classifier for (2) must
\emph{learn} that \InSlice is a strong predictor.
%
In contrast, a classifier for (3) must only learn about the syntactic
features, the decision to discard samples where \InSlice is disabled has
already been made by a human.
%
This has a few additional advantages: it reduces the set of candidate
locations by a factor of 7 on average, and it guarantees that any
prediction made by the classifier can fix the type error.
%
We expect that (2) will perform better than (1) as it contains more
information, and that (3) will perform better than (2) as the classifier
does not have to learn the importance of \InSlice.

@@

We investigate the relative impact of the other
three classes of features discussed in \autoref{sec:features}, assuming
we have discarded expressions not in the type error slice.
%
For this experiment we consider again a baseline of only local syntactic
features, extended by each combination of
%
(1) expression size;
(2) contextual syntactic features; and
(3) typing features.
%
As before, we perform a 10-fold cross-validation with $\eta = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs to make the differences more apparent.
%
% \begin{table}[ht]
%   \caption{
%     Impact of Contextual Features on Accuracy.
%     \ES{TODO: load these numbers from CSV}
%   }\label{tab:contextual-features}
%   \centering
%   \begin{tabular}{lrcrrrrcrrrr}
%     \toprule
%                              &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
%                                                \cmidrule{4-7}                        \cmidrule{9-12}
%     Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
%     \midrule
%     Local Syntax             &  44         & & 56.6\% & 72.6\% & 83.2\% & 57.7\% & & 57.6\% & 72.6\% & 83.1\% & 57.7\% \\
%     \midrule
%     + Size                   &  45         & & 57.2\% & 73.7\% & 83.0\% & 57.4\% & & 60.9\% & 75.5\% & 84.1\% & 57.9\% \\
%     + Context                & 220         & & 61.1\% & 78.7\% & 86.7\% & 63.0\% & & 71.5\% & 84.8\% & 90.8\% & 69.0\% \\
%     + Types                  & 102         & & 63.1\% & 77.8\% & 85.2\% & 61.6\% & & 73.4\% & 85.4\% & 90.7\% & 68.9\% \\
%     \midrule
%     + Context + Size         & 221         & & 61.1\% & 78.8\% & 86.3\% & 62.4\% & & 71.5\% & 84.4\% & 90.8\% & 69.0\% \\
%     + Types + Size           & 103         & & 61.9\% & 78.7\% & 85.6\% & 62.0\% & & 73.1\% & 85.2\% & 91.1\% & 69.5\% \\
%     + Context + Types        & 275         & & 63.1\% & 80.9\% & 88.0\% & 65.0\% & & 77.2\% & 88.3\% & 92.5\% & 72.6\% \\
%     \midrule
%     + Context + Types + Size & 276         & & 62.8\% & 80.5\% & 88.1\% & 65.1\% & & 77.3\% & 88.1\% & 92.7\% & 72.7\% \\
%     \bottomrule
%   \end{tabular}
%   % \begin{minipage}{0.49\linewidth}
%   % \centering
%   % \hiddenF
%   % \begin{tabular}{lrrrr}
%   %   \toprule
%   %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
%   %   \midrule
%   %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
%   %   \midrule
%   %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
%   %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
%   %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
%   %   \midrule
%   %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
%   %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
%   %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
%   %   \midrule
%   %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
%   %   \bottomrule
%   % \end{tabular}
%   % \end{minipage}
% \end{table}
%
\paragraph{Results}
\autoref{fig:context-utility} summarizes the results of this experiment.
%
The \linear classifier and the \hiddenFH start off
competitive when given only local syntactic features, but the \hiddenFH
quickly outperforms as more features are added.

>>>>

\subsubsection{Type Error Slice}
\label{sec:type-error-slice}
First we must justify our decision to automatically exclude expressions
outside the minimal type error slice from consideration.
%
% The \InSlice feature should be highly predictive --- a fix must change
% at least one expression in the type error slice.
% %
% Thus, our first experiment seeks to quantify the impact of \InSlice by
% comparing the accuracy of our classifiers on three sets of features:
%
Thus, we compare our classifiers on three sets of features:
%
\begin{enumerate}
\item A baseline with only local syntactic features and no
  preemptive filtering by \InSlice.
\item The features of (1) extended with \InSlice.
\item The same features as (1), but we preemptively discard samples
  where \InSlice is disabled.
\end{enumerate}
%
The key difference between (2) and (3) is that a classifier for (2) must
\emph{learn} that \InSlice is a strong predictor.
%
In contrast, a classifier for (3) must only learn about the syntactic
features, the decision to discard samples where \InSlice is disabled has
already been made by a human.
%
This has a few additional advantages: it reduces the set of candidate
locations by a factor of 7 on average, and it guarantees that any
prediction made by the classifier can fix the type error.
%
We expect that (2) will perform better than (1) as it contains more
information, and that (3) will perform better than (2) as the classifier
does not have to learn the importance of \InSlice.

@@

We investigate the relative impact of the other
three classes of features discussed in \autoref{sec:features}, assuming
we have discarded expressions not in the type error slice.
%
For this experiment we consider again a baseline of only local syntactic
features, extended by each combination of
%
(1) expression size;
(2) contextual syntactic features; and
(3) typing features.
%
As before, we perform a 10-fold cross-validation with $\eta = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs to make the differences more apparent.
%
% \begin{table}[ht]
%   \caption{
%     Impact of Contextual Features on Accuracy.
%     \ES{TODO: load these numbers from CSV}
%   }\label{tab:contextual-features}
%   \centering
%   \begin{tabular}{lrcrrrrcrrrr}
%     \toprule
%                              &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
%                                                \cmidrule{4-7}                        \cmidrule{9-12}
%     Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
%     \midrule
%     Local Syntax             &  44         & & 56.6\% & 72.6\% & 83.2\% & 57.7\% & & 57.6\% & 72.6\% & 83.1\% & 57.7\% \\
%     \midrule
%     + Size                   &  45         & & 57.2\% & 73.7\% & 83.0\% & 57.4\% & & 60.9\% & 75.5\% & 84.1\% & 57.9\% \\
%     + Context                & 220         & & 61.1\% & 78.7\% & 86.7\% & 63.0\% & & 71.5\% & 84.8\% & 90.8\% & 69.0\% \\
%     + Types                  & 102         & & 63.1\% & 77.8\% & 85.2\% & 61.6\% & & 73.4\% & 85.4\% & 90.7\% & 68.9\% \\
%     \midrule
%     + Context + Size         & 221         & & 61.1\% & 78.8\% & 86.3\% & 62.4\% & & 71.5\% & 84.4\% & 90.8\% & 69.0\% \\
%     + Types + Size           & 103         & & 61.9\% & 78.7\% & 85.6\% & 62.0\% & & 73.1\% & 85.2\% & 91.1\% & 69.5\% \\
%     + Context + Types        & 275         & & 63.1\% & 80.9\% & 88.0\% & 65.0\% & & 77.2\% & 88.3\% & 92.5\% & 72.6\% \\
%     \midrule
%     + Context + Types + Size & 276         & & 62.8\% & 80.5\% & 88.1\% & 65.1\% & & 77.3\% & 88.1\% & 92.7\% & 72.7\% \\
%     \bottomrule
%   \end{tabular}
%   % \begin{minipage}{0.49\linewidth}
%   % \centering
%   % \hiddenF
%   % \begin{tabular}{lrrrr}
%   %   \toprule
%   %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
%   %   \midrule
%   %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
%   %   \midrule
%   %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
%   %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
%   %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
%   %   \midrule
%   %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
%   %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
%   %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
%   %   \midrule
%   %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
%   %   \bottomrule
%   % \end{tabular}
%   % \end{minipage}
% \end{table}
%
\paragraph{Results}
\autoref{fig:context-utility} summarizes the results of this experiment.
%
The \linear classifier and the \hiddenFH start off
competitive when given only local syntactic features, but the \hiddenFH
quickly outperforms as we add features.

### 2017-04-11 paper/oopsla17-submission/gallery.tex
fiddling with graphs instead of tables..
use colors in opening example
prune some of the more obnoxious orphans

\definecolor{tree}{HTML}{66C2A5}
\definecolor{sherrloc}{HTML}{FC8D62}
\definecolor{fix}{HTML}{8DA0CB}

@@

\paragraph{Currying Considered Harmful?}
% data/sp14/1204.ml
Our next example is another ill-fated attempt at |clone|.
%
\begin{ecode}
  let rec clone x n =
    let rec loop (*@\colorbox{fix!75}{x n acc}@*) =
      if n < 0 then
        acc
      else
        (*@\colorbox{sherrloc!75}{loop} \colorbox{tree!75}{(x, (n - 1), (x :: acc))}@*) in
    loop (x, n, [])
\end{ecode}
%
The issue here is that \ocaml functions are \emph{curried} by default
--- \ie they take their arguments one at a time --- but our student has
called the inner |loop| with all three arguments in a tuple.
%
Many experienced functional programmers would choose to keep |loop|
curried and rewrite the calls, however our student decides instead to
\emph{uncurry} |loop|, making it take a tuple of arguments.
%
\sherrloc blames the recursive call to |loop| while our classifier
blames the tuple of arguments --- a reasonable suggestion, but alas not
the answer the student was looking for.

@@

\paragraph{Extracting the Digits of an Integer}
% data/sp14/2176.ml
Consider first a simple recursive function |digitsOfInt| that extracts
the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ (*@\colorbox{tree!75}{[ \colorbox{sherrloc!75}{digitsOfInt (n / 10)} ]}@*)
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
actually correct!

@@

Our \dectree correctly predicts that the recursive call blamed by
\sherrloc should not be blamed; a similar examination suggests that the
crucial observation is that the parent of the recursive call is a data
constructor application.
%

>>>>

@@

\paragraph{Currying Considered Harmful?}
% data/sp14/1204.ml
Our next example is another ill-fated attempt at |clone|.
%
\begin{ecode}
  let rec clone x n =
    let rec loop (*@\colorbox{fix!75}{x n acc}@*) =
      if n < 0 then
        acc
      else
        (*@\colorbox{sherrloc!75}{loop} \colorbox{tree!75}{(x, (n - 1), (x :: acc))}@*) in
    loop (x, n, [])
\end{ecode}
%
The issue here is that \ocaml functions are \emph{curried} by default
--- \ie they take their arguments one at a time --- but our student has
called the inner |loop| with all three arguments in a tuple.
%
Many experienced functional programmers would choose to keep |loop|
curried and rewrite the calls, however our student decides instead to
\emph{uncurry} |loop|, making it take a tuple of arguments.
%
\sherrloc blames the recursive call to |loop| while our classifier
blames the tuple of arguments --- a reasonable suggestion, but not
the answer the student expected.

@@

\paragraph{Extracting the Digits of an Integer}
% data/sp14/2176.ml
Consider first a simple recursive function |digitsOfInt| that extracts
the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ (*@\colorbox{tree!75}{[ \colorbox{sherrloc!75}{digitsOfInt (n / 10)} ]}@*)
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
correct!

@@

Our \dectree correctly predicts that the recursive call blamed by
\sherrloc should not be blamed; a similar examination suggests that the
crucial observation is that the recursive call's parent is a data
constructor application.
%
